\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-07 paper/draft.tex
Complete pash on evaluation

\heading{Programs}
In a recent celebration of \unix's 50-year legacy, Bell Labs created 37 challenges~\cite{unix50} solvable by means of composing \unix pipelines.
We found unofficial solutions to all-but-three problems on GitHub~\cite{unix50sol}, expressed as pipelines with 2--12 stages (avg.: 5.58).
We consider this a good set of benchmarks because
  (i) the problems were designed to highlight \unix's modular philosophy~\cite{mcilroy1978unix} and make extensive use of standard commands under a variety of flags, and
  (ii) the solutions were written by non-experts (contrary to \S\ref{ours}).
% ---stressing a patterns that we, \sys's authors, did not
% necessarily anticipate \kk{I would change that to ``stressing a
%   variety of patterns'' since it doesn't really matter if we
%   anticipated them (especially since we can address any issues and
%   then re-run them).} . 
\sys executes each pipeline as-is; 
  if an obvious fix improves performance, we report the speedup of both versions. %  original and modified. %  pipeline.
% TODO: talk about correct wrt developer intentions vs. correct wrt problem definition

@@

%% \TODO{An important point that needs to be made clear is that these are
%%   taken from the wild unmodified. And that Pash never leads to
%%   slowdown (except for the trivial head cases, where the slowdown is
%%   neligible).}
\heading{Results}
\sys's parallelism is set to $16\times$, as many of these pipelines are relatively long (\S\ref{eval}, Parallelism).
Their input was generated by multiplying their original input several times, enough to reach 10GB.
Fig.~\ref{fig:unix50-individual} shows the speedup (left) over the sequential runtime (right) for all 34 pipelines. We use the pipeline index to refer to them in both plots and discussion.
%% (\nv{Pipelines are referred by number, in both the plot and the discussion below.})
Average speedup is \todo{$6.02\times$}, median is $6.48\times$, and weighted average (with the absolute times as weights) is $5.75\times$.
 % the \todo{geometric mean is $1.08\times$},
We do not include three pipelines that take less than 0.1 seconds to execute because they only output the first line of their input using \ttt{head}.

@@

\heading{Results}
The complete pipeline executes in 44m2s for five years (82GB) of data.
\sys with 2/10$\times$ parallelism leads to 1.86/2.44$\times$ speedup, with different phases seeing different benefits:
  1.77/1.99$\times$ (vs. 33m58s) for all the pre-processing and 
  2.30/10.79$\times$ speedup (vs. 10m4s) for computing the max.
(For why speedup is higher than parallelism, see \S\ref{ours}.) % 's Discussion.

@@

\heading{Take-aways}
\sys can be applied to programs of notable size and input
%% \todo{larger} % real/realistic? implies that prior benchmarks were not  
%% programs
%% with modern input sides
to offer significant acceleration.
A broader take-away is that \sys is also able to extract parallelism from fragments that are not purely compute-intensive regions, which are the focus of conventional parallelization systems.

@@

\heading{Results}
% To simplify execution (and avoid Wikipedia), we have saved a recent Wikipedia version on a server within the network and to redirect all \ttt{curl} requests there.
%% As the \todo{first} component contains a cycle (URLs $\rightarrow$ HTML $\rightarrow$ URLs), \sys's conservative transformations do not attempt to parallelize it.
The original script takes 191min to execute on 1\% of Wikipedia (1.3GB).
With 2/16$\times$-parallelism, \sys brings it down to 97/15min ($1.97/12.7\times$), with the majority speedup coming from the HTML-to-text conversion. % which uses \ttt{pandoc}, a utility written in Haskell.

@@

\pichskip{15pt}% Horizontal gap between picture and text
\parpic[r][t]{%
  \begin{minipage}{40mm}
    \includegraphics[width=\linewidth]{\detokenize{./figs/sort_baseline_comparison_scaleup.pdf}}%
    %% \captionof{figure}{This is a tiger.}
  \end{minipage}
}
\heading{Parallel Sort}
%% \kk{Explain what is the purpose of the
%%   microbenchmark. We want to establish that \sys scalability is as
%%   good as it could be.}
We first compare a single GNU \ttt{sort} optimized by \sys ($S_{PaSh}$) versus the same \ttt{sort} with the 
%% \begin{wrapfigure}[10]{r}{0.25\textwidth}
%%   \centering
%%   %% \begin{center}
%%     \includegraphics[width=0.25\textwidth]{\detokenize{./figs/sort_baseline_comparison_scaleup.pdf}}
%%   %% \end{center}
%% %   \vspace{-15pt}
%% % \caption{ \textbf{Comparison with \ttt{sort -{}-parallel}.} }
%% \end{wrapfigure}
\ttt{--parallel} flag set ($S_{GNU}$). % , shown on the side.
% The results are shown on the side-figure below.
While \ttt{--parallel} is not a general solution, the comparison serves to establish a baseline for \sys.
$S_{GNU}$'s parallelism is configured to $2\times$ that of $S_{PaSh}$'s (\ie the rightmost plot point for $S_{GNU}$ is for \ttt{--parallelism=127}), 
to account for \sys's additional merge processes.
% and as such we note that the measurements are harsh for \sys.

A few points are worth noting.
$S_{PaSh}$ without \ttt{eager} performs comparably to $S_{GNU}$, and with \ttt{eager} it outperforms $S_{GNU}$ ($\sim2\times$);
  this is because \ttt{eager} adds intermediate buffers %(on disk)
  between merge phases.
%% On the other hand, \ttt{sort --parallel} sorts the input trying to
%% minimize additional memory (so doesn't allocate intermediate
%% buffers).
%% %% \kk{If we want, we can add here numbers where we allow sort --parallel
%% %%   to use additional ram for intermediate buffers (using the -S flag).}
%% Of course, it would be possible to manually create a longer pipeline
%% and use intermediate buffers to speed this up, but this defeats the
%% purpose of having a simple solution.
$S_{GNU}$ indicates that \ttt{sort}'s scalability is inherently limited (\ie due to \ttt{sort} and not \sys);
  this is why all scripts that contain \ttt{sort} (\eg \S\ref{ours}--\ref{macro2}) do not go above $8\times$.
Finally, the comparison also shows \sys's benefits to command developers:
  a low-effort parallelizability annotation allows similar or better scalability than a custom flag added by developers.

@@

There are a few possible ways one might attempt to use GNU \ttt{parallel} on this program.
They could use it on the 8th stage, assuming they know it is a bottleneck, bringing execution down to 304.4s ($1.8\times$ speedup).
Alternatively, they could (incorrectly) sprinkle \ttt{parallel} across the entire program---a strategy simplified by \ttt{parallel}'s \ttt{stdin} redirection feature.
This would lead to 3.2$\times$ performance improvements but severely incorrect results with respect to the sequential execution---with 92\% of the output showing a difference between sequential and parallel execution.
\sys's conservative program transformations will not operate on fragments with unclear parallelizability properties.
	% or commands whose inputs depend on earlier outputs.
% GNU \ttt{parallel} is oblivious to command semantics, risking breaking program semantics as in this case.

>>>>

\heading{Programs}
In a recent celebration of \unix's 50-year legacy, Bell Labs created 37 challenges~\cite{unix50} solvable by means of composing \unix pipelines.
We found unofficial solutions to all-but-three problems on GitHub~\cite{unix50sol}, expressed as pipelines with 2--12 stages (avg.: 5.58).
We consider this a good set of benchmarks because
  (i) the problems were designed to highlight \unix's modular philosophy~\cite{mcilroy1978unix}, 
  (ii) they make extensive use of standard commands under a variety of flags, and
  (iii) the solutions were written by non-experts (contrary to \S\ref{ours}).
% ---stressing a patterns that we, \sys's authors, did not
% necessarily anticipate \kk{I would change that to ``stressing a
%   variety of patterns'' since it doesn't really matter if we
%   anticipated them (especially since we can address any issues and
%   then re-run them).} . 
\sys executes each pipeline as-is, without any modification.
  % if an obvious fix improves performance, we report the speedup of both versions. %  original and modified. %  pipeline.

@@

\heading{Results}
\sys's parallelism is set to $16\times$, as many of these pipelines are relatively long (\S\ref{eval}, Parallelism).
Their input was generated by multiplying their original input several times, enough to reach 10GB.
Fig.~\ref{fig:unix50-individual} shows the speedup (left) over the sequential runtime (right) for all 34 pipelines. We use the pipeline index to refer to them in both plots and discussion.
%% (\nv{Pipelines are referred by number, in both the plot and the discussion below.})
Average speedup is \todo{$6.02\times$}, median is $6.48\times$, and weighted average (with the absolute times as weights) is $5.75\times$.
 % the \todo{geometric mean is $1.08\times$},
We do not include three pipelines that take less than 0.1 seconds to execute because they only output the first line of their input using \ttt{head}.

@@

\heading{Results}
The complete pipeline executes in 44m2s for five years (82GB) of data.
\sys with 2/10$\times$ parallelism leads to 1.86/2.44$\times$ speedup, with different phases seeing different benefits:
  1.77/1.99$\times$ (vs. 33m58s) for all the pre-processing and 
  2.30/10.79$\times$ speedup (vs. 10m4s) for computing the max.
The reason why speedup is higher than parallelism is also also explained in the Discussion of \S\ref{ours}.

@@

\heading{Take-aways}
\sys can be applied to programs of notable size and input
%% \todo{larger} % real/realistic? implies that prior benchmarks were not  
%% programs
%% with modern input sides
to offer significant acceleration.
A broader take-away is that \sys is also able to extract parallelism from fragments that are not purely compute-intensive regions, \ie the usual focus of conventional parallelization systems.

@@

\heading{Results}
% To simplify execution (and avoid Wikipedia), we have saved a recent Wikipedia version on a server within the network and to redirect all \ttt{curl} requests there.
%% As the \todo{first} component contains a cycle (URLs $\rightarrow$ HTML $\rightarrow$ URLs), \sys's conservative transformations do not attempt to parallelize it.
The original script takes 191min to execute on 1\% of Wikipedia (1.3GB).
With 2/16$\times$-parallelism, \sys brings it down to 97/15min ($1.97/12.7\times$), with the majority of the speedup coming from the HTML-to-text conversion. % which uses \ttt{pandoc}, a utility written in Haskell.

@@

\pichskip{15pt}% Horizontal gap between picture and text
\parpic[r][t]{%
  \begin{minipage}{40mm}
    \includegraphics[width=\linewidth]{\detokenize{./figs/sort_baseline_comparison_scaleup.pdf}}%
    %% \captionof{figure}{This is a tiger.}
  \end{minipage}
}
\heading{Parallel Sort}
%% \kk{Explain what is the purpose of the
%%   microbenchmark. We want to establish that \sys scalability is as
%%   good as it could be.}
We first compare a single GNU \ttt{sort} parallelized by \sys ($S_{PaSh}$) versus the same \ttt{sort} with the 
%% \begin{wrapfigure}[10]{r}{0.25\textwidth}
%%   \centering
%%   %% \begin{center}
%%     \includegraphics[width=0.25\textwidth]{\detokenize{./figs/sort_baseline_comparison_scaleup.pdf}}
%%   %% \end{center}
%% %   \vspace{-15pt}
%% % \caption{ \textbf{Comparison with \ttt{sort -{}-parallel}.} }
%% \end{wrapfigure}
\ttt{--parallel} flag set ($S_{GNU}$). % , shown on the side.
% The results are shown on the side-figure below.
While \ttt{--parallel} is not a general solution, the comparison serves to establish a baseline for \sys.
$S_{GNU}$'s parallelism is configured to $2\times$ that of $S_{PaSh}$'s (\ie the rightmost plot point for $S_{GNU}$ is for \ttt{--parallelism=127}), 
to account for \sys's additional runtime processes.
% and as such we note that the measurements are harsh for \sys.

A few points are worth noting.
$S_{PaSh}$ without \ttt{eager} performs comparably to $S_{GNU}$, and with \ttt{eager} it outperforms $S_{GNU}$ ($\sim2\times$);
  this is because \ttt{eager} adds intermediate buffers %(on disk)
  that ensure CPU utilization is high.
%% On the other hand, \ttt{sort --parallel} sorts the input trying to
%% minimize additional memory (so doesn't allocate intermediate
%% buffers).
%% %% \kk{If we want, we can add here numbers where we allow sort --parallel
%% %%   to use additional ram for intermediate buffers (using the -S flag).}
%% Of course, it would be possible to manually create a longer pipeline
%% and use intermediate buffers to speed this up, but this defeats the
%% purpose of having a simple solution.
$S_{GNU}$ indicates that \ttt{sort}'s scalability is inherently limited (\ie due to \ttt{sort} and not \sys);
  this is why all scripts that contain \ttt{sort} (\eg \S\ref{ours}--\ref{macro2}) do not go above $8\times$.
Finally, the comparison also shows \sys's benefits to command developers:
  a low-effort parallelizability annotation achieves better scalability than a custom flag (and underlying parallel implementation) added by developers.

@@

There are a few possible ways one might attempt to use GNU \ttt{parallel} on this program.
They could use it on the 8th stage, assuming they know it is a bottleneck, bringing execution down to 304.4s ($1.8\times$ speedup).
Alternatively, they could (incorrectly) sprinkle \ttt{parallel} across the entire program---a strategy simplified by \ttt{parallel}'s \ttt{stdin} redirection feature.
This would lead to 3.2$\times$ performance improvements but severely incorrect results with respect to the sequential execution---with 92\% of the output showing a difference between sequential and parallel execution.
\sys's conservative program transformations are not applied in program fragments with unclear parallelizability properties.
	% or commands whose inputs depend on earlier outputs.
% GNU \ttt{parallel} is oblivious to command semantics, risking breaking program semantics as in this case.
